### 2013-07-08 paper.tex

\section{Introduction}
Tree volume measurement is one of the many parameters in documenting and profiling trees and is at the center of forest inventory.Estimates of tree volume are important for planning timber harvests and modelling carbon sequestration. The need for information about Earth's natural resources, including tree volume, has been increasing exponentially in recent times. To be able to accomplish this in forestry, it is a necessity to be able to collect large amounts of data easily and on a regular basis.

Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}
Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches do not come without their downfalls.

@@

SilviaTerra, a high-tech startup company, however is willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results. This is achieved through computer exactitude combined with statistical analysis. Recently a new project was started, with this research project developing the mathematical back end, to allow users to easily get a correct volume measurement by only using their smart phone.This project aims to solve the problem of obtaining an accurate volume reading of tree, cheaply and non-destructively. This will be done by using a combination of photos taken of the tree, various sensor data from the phone and mathematical computation.

@@

The next stage is retrieving the real world height to each set pixel coordinates as well as the width between them, which will be the diameter of the tree. The first step taken is to get the height. The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view. This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane on the camera sensor. With this method it is possible to find out the angle to each pixel coordinate, example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also calculated so that the width can later be determined.

The final part of obataining the parameters for the set of pixel coordinates at that height is to calculate the width between them. This is accomplished by employing the method for calculating the distance to the tree, in reverse. In this case what is known in figure \ref{horizontal_triangle} is the distance $d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$. $W_2$ is calculated by multiplying $d$ with $\tan{\varphi}$. $W_2$ is then again multiplied by 2 and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement. To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ dispensing the physical width of the tree at that certain height.

\subsection{Volume}
Now the only thing left to produce the final result is to use these newly calculated widths and heights to obtain the total volume. The current implemented method of doing this is by integrating all the points using the trapezoidal rule. All the widths are divided by 2 to get the radius. Then they are squared and multiplied with $\pi$ to the obtain the basal area at each height. The basal area is plotted along with their corresponding heights. The trapezoidal rule then defines that the area between two points is: $\frac{a_1 + a_2}{2} \times h$. The area between all the points is then summed together.

@@

\section{Discussion}
\subsection{Error tolerance}
As seen in table \ref{heights} either when the vertical tilt angle is very wide, standing close to the tree, or very narrow, standing far away, leads to error. It's not clear what is causing this but there a two possibl reasons. The first one is that because of the short focal length of the camera and overall poor quality when compared to a real camera, causes various forms of distrotion, especially around the edges of the image. When the person is standing close to the tree so the vertical tilt angle is over 45\degree, the points tend to lie exactly there, around the edges of the image. The distortion then affects these points causing the pixel measurements to be incorrect. The other problem occurs in the direct opposite situation, when the person is standing far away and the angle is very narrow. Naturally what happens is that the resolution of the camera just isn't high enough and the pixel measurements just aren't accurate anymore. However most of this distortion seems to only affect the vertical measurements because when looking at table \ref{widths} it is clearly seen that the readings do not stray much at all from their actual values. Despite this the deviation in the height readings still alters the volume calculations. Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which is basically means they should stand somewhere between one tree length to two tree length's from the tree.
\subsection{Limitations}
As pointed out in the previous section the main hindrance attatched to this application is the actual camera. The problem is that most mobile phone cameras have a lot of different distortions that make it hard to make accurate measurements. Also when the tree is far away, the resolution of the camera can have a large affect on the final result. Human error is still large factor. If the person is not holding the phone properly, takes a blurry photo or in any other way alters the image, will cause the calculations to be incorrect. Mechanical Turk is also a weak link because it is essentially humans doing the work and there nothing stopping them from also making mistakes. 
\subsection{Future development}
The main area that could be expanded on is the edge detection. As explained in the background, today computer vision is still far from perfect. However that will hopefully change in the near future and things like Mechanical Turk can be dropped for a more efficient system that produces little to no error and works autonomously. If camera technology in mobile phones doesn't improve then being able to correct the various distortions would also make an improvement. The possibilty might also be there in the future to have the application calculate extra things like canopy area and even adding in the volume of external branches to the main volume.

>>>>

\section{Introduction}
Tree volume measurement is one of the many parameters used when documenting and profiling trees and is at the center of forest inventory.Estimates of tree volume are important for planning timber harvests and modelling carbon sequestration. The need for information about Earth's natural resources, including tree volume, has been increasing exponentially in recent times. To be able to accomplish this in forestry, it is a necessity to be able to collect large amounts of data easily and on a regular basis.

Unfortunately estimating tree volume is expensive, time consuming and in most cases the data collected doesn't meet today's standards \cite{digital imaged based tree measurement for forest inventory}
Most methods used today to measure individual trees are often done manually with a variety of tools. However the use of these analogue approaches does not come without it's downfalls.

@@

SilviaTerra, a high-tech startup company,is willing to change that and have already developed several tools for scanning, analyzing and mapping areas of forestry to give accurate results. This is achieved through computer exactitude combined with statistical analysis. Recently a new project was started, with this research project developing the mathematical back end, to allow users to easily get a correct volume measurement by only using their smart phone.This project aims to solve the problem of obtaining an accurate volume reading of tree, cheaply and non-destructively. This will be done by using a combination of photos taken of the tree, various sensor data from the phone and mathematical computation.

@@

The next stage is retrieving the real world height to each set of pixel coordinates as well as the width between them, which will be the diameter of the tree. The first step taken is to get the height. The vertical image plate scale, which is the ratio between degrees and pixels in the vertical axis, is calculated by dividing the vertical resolution of the image with the vertical angle of view. This can be done because it is assumed that the silhouette of the tree is on a plane in the real world which corresponds to the image plane on the camera sensor. With this method it is possible to find out the angle to each pixel coordinate, for example $P$ in figure \ref{triangle1} and \ref{triangle2}, in the angle of view. This is angle is denoted by $\theta$ in both figure \ref{triangle1} and \ref{triangle2}. However if the pixel coordinate was identified in the bole image, when the phone was tilted upwards, then angle between the distance to the tree and the lower boundary of the vertical angle of view, $B$, will need to be added, which is represented by $\varphi$ in figure \ref{triangle2}. The diagonal distance to $P$ is also calculated so that the width can later be determined.

The final part of obataining the parameters for each set of pixel coordinates is to calculate the width between them. This is accomplished by employing the method for calculating the distance to the tree, in reverse. In this case what is known in figure \ref{horizontal_triangle} is the distance $d$ and the angle $\theta$ but not the width of the tree, $X_1$ to $X_2$. $W_2$ is calculated by multiplying $d$ with $\tan{\varphi}$. $W_2$ is then again multiplied by 2 and then divided by the number of pixels in the horizontal resolution of the image, once more returning the ratio between pixels and the corresponding real world measurement. To finish off this ratio is the multiplied by the number of pixels between $X_1$ and $X_2$ resulting in the physical width of the tree at that certain height.

\subsection{Volume}
Now the only thing left to produce the final result is to use these newly calculated widths and heights to obtain the total volume. The new calculated heights and lengths are used to calculate a loess curve. A loess curve is a line that is fitted to the input points through local regression. The loess curve is then used to produce a polynomial spline function that can then be used to calculate the diameter at very small height intervals. At each new interval the volume can be calculated and is then added onto the current total volume. This will result in a much more accurate reading than by just numerically intergrating the original points.

@@

\section{Discussion}
\subsection{Error tolerance}
As seen in table \ref{heights} either when the vertical tilt angle is very wide, standing close to the tree, or very narrow, standing far away, leads to error. It's not clear what is causing this but there a two possible reasons. The first one is that because of the short focal length of the camera and overall poor quality when compared to a real camera, causes various forms of distrotion, especially around the edges of the image. When the person is standing close to the tree so the vertical tilt angle is over 45\degree, the points tend to lie exactly there, around the edges of the image. The distortion then affects these points causing the pixel measurements to be incorrect. The other problem occurs in the direct opposite situation, when the person is standing far away and the angle is very narrow. Naturally what happens is that the resolution of the camera just isn't high enough and the pixel measurements just aren't accurate anymore. However most of this distortion seems to only affect the vertical measurements because when looking at table \ref{widths} it is clearly seen that the readings do not stray much at all from their actual values. Despite this the deviation in the height readings still alters the volume calculations. Ideally a person should stand so that the vertical tilt angle lies between 45\degree and 25\degree which basically means they should stand somewhere between one tree length to two tree length's from the actual tree.
\subsection{Limitations}
As pointed out in the previous section the main hindrance attatched to this application is the actual camera. The problem is that most mobile phone cameras have a lot of different distortions that make it hard to make accurate measurements. Also when the tree is far away, the resolution of the camera can have a large affect on the final result. Human error is still large factor. If the person is not holding the phone properly, takes a blurry photo or in any other way alters the image, will cause the calculations to be incorrect. Mechanical Turk is also a weak link because it is essentially humans doing the work and there is nothing stopping them from also making mistakes. 
\subsection{Future development}
The main area that could be expanded on is the edge detection. As explained in the background, today computer vision is still far from perfect. However that will hopefully change in the near future and things like Mechanical Turk can be dropped for a more efficient system that produces little to no error and works autonomously. If camera technology in mobile phones doesn't improve then being able to correct the various distortions would also make an improvement. The possibilty might also exist in the future to have the application calculate extra things like canopy area and even adding in the volume of external branches to the main volume.
